FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-02-28 ess_marmo.tex

\section{Discussion and Perspectives}
\label{sec:disc}
FITS discussion in astronomy\citep{THOMAS2015133} 
let's astronomy learn from planetary sciences too: accent must be put on interoperability
.... blabla
Planetary surface is lacking open and general software for massive processing and can benefit
of astronomy experience. In addition to that WCS projection scheme is simpler than the
historical Earth-based one used in planetary applications. 
Archiving flexibility and data model are today approached by PDS4 and EPN-TAP efforts.
Reducing custom data format conversions during the processing chain but
allowing easy conversion for final products following the needed application.

>>>>

\section{Discussion and Perspectives}
\label{sec:disc}
An extension to the FITS standard has been proposed to encompass planetary
surface data description.
Planetary surface investigation is lacking open and general software for massive data
processing and can benefit from astronomy experience in this field.

We are aware that FITS is today subject of questioning in astronomical applications
\citep{THOMAS2015133}.
We are not convinced that one format can solve all the issues faced from long-term
archiving to on-line streaming.
Planetary science research community has used a large variety of format customizing
specific software.
We want to stress here that using FITS could be an efficient compromise, reducing custom
data format conversions during the processing chain and
allowing easy conversion for final products following the needed application.

FITS as an open and flexible data and metadata exchange and archiving format, could be
inserted in the PDS4 data model scheme in order to address the issue of more complex
metadata and data relationships.

In order to improve interoperability and knowledge sharing between astronomy and
planetary science, we will soon finalize the integration of FITS specifications in the
GDAL library.
In the framework of the VESPA project a review of the available and planned mission data
archives is ongoing in order to evaluate their insertion in the planetary VO structure
with a possible on-demand conversion in FITS format.
